Deriving an Executable Model of a Data Stream Protocol from its Specification

### 2009-10-12 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
ModelBasedTestCaseGenerationDataStreamProtocol paper: first almost done draft finished, everything smoothed out, only conclusions are missing

%\newcommand{\notev}[1]{\xspace$\textcolor{blue}{\omega^\textsf{vieri}}$\marginpar{\scriptsize\textsf{Vieri:} #1}}
\newcommand{\notev}[1]{\xspace\marginpar{\scriptsize\textsf{Vieri:} #1}}
\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:} #1}}
\newcommand{\noted}[1]{\xspace$\textcolor{red}{\dagger^\textsf{dragan}}$\marginpar{\scriptsize\textsf{Dragan:} #1}}

@@

%---------------------------------------------------------------------
% New commands, macros, \etc
%---------------------------------------------------------------------

@@

%=====================================================================

@@

\author{Joseph R. Kiniry, Vieri del Bianco\\
UCD CASL: Complex and Adaptive Systems Laboratory and\\
School of Computer Science and Informatics,\\
University College Dublin,\\
Belfield, Dublin 4, Ireland,\\
kiniry@acm.org, vieri.delbianco@gmail.com\\
}

@@

%======================================================================
%\thispagestyle{empty}
\begin{abstract}

\begin{verbatim}

OBJECTIVE: 
Specifications used as a blueprint for system development.
Specifications to generate test suite to validate the developed system.
Specifications to validate simulator of the developed system.
Apply to data stream processing: verify the stream is parsed correctly, 
verify the parsed stream output stream can be simulated.

METHODS: 
JML model.
JML tools to generate JUnit test suite.
JUnit test suite applied to developed system, to verify its correctness.
JUnit test suite applied to simulator, to verify its correctness.

RESULTS: 
Proof developed system same behavior as simulator.
Proof developed system and simulator verify the specifications.

CONCLUSIONS:
Specifications serves two purposes, and indirectly a third one.

\end{verbatim}

  Sensor equipped devices typically communicate with a streaming
  packet-based network protocol.  The communication is focused on
  controlling the sensors and report the sensors measurements.  The
  \STSB is an example of such a device, it communicates with an
  arbitrary computer system, the \STPU (running Linux or Windows, at
  the moment) over a USB bus using a streaming packet-based protocol.
  The \STU is a custom-designed, but inexpensive, sensor platform: it
  is composed by a one or more \STSBs paired with a general-purpose
  small-scale compute node, the \STPU (including everything from PDAs
  to powerful touchscreen-based portable computers).  There are over a
  dozen sensors on the \STSB itself, and new sensors can be added
  either using expansion slots on the \STSB, or adding them directly
  to the USB bus.  A third-party low-level library, provided by the
  USB chip vendor, the USB chip that has been used in the \STSB,
  creates and parses generic USB communication packets.

  A custom medium-level Java driver creates and parses the \STSB
  streaming communication packets.  The driver is formally specified,
  validated and verified with rigorous unit testing.  Because of this
  process and combination of technologies, using this specification we
  are able to both communicate with a real \ST \textbf{and} generate a
  formally specified, executable, functional model-based specification
  for the protocol itself written in the Java Modeling Language.
  \notev{executable? how? what does this mean?}  This protocol model
  is itself an executable stream simulator that can be used to develop
  new \ST user applications without having the \ST hardware at all.
  Consequently, from this single specification, we can test that the
  \ST hardware communicates as expected, test that our \ST simulator
  communicates as expected, and simulate the \ST protocol stream
  itself.

@@

%======================================================================
\section{Introduction}

@@

Because of hard time constraints, and the initial unavailability of the 
custom board, we needed to progress concurrently with all the development tasks: 
the specification of the communication protocol with the \STSB, the development of 
the controlling code in the \STSB, the development of the driver to communicate 
with the \STSB, and the development of the simulators needed for future tests 
of the system (the controlling code of the board was developed by the external 
manufacturer of the board, hence, it is not taken into account here).

\todo{explain and refine the problem}

\todo{references: search other approaches to similar problems}

To tackle the complexity of the concurrent tasks the formal model of the protocol 
was used as a focus point that led all the other development tasks. 
The formal model of the protocol was used to check at run time the correctness of 
the /SB, the driver, the simulator, and the hand made test suites of the driver 
and the simulator.
The formal model was used to automatically generate a test suite to verify that
both the driver and the simulator were compliant to the specification, and eventually 
to verify the specification itself.

The approach proved itself effective, at the point that, when the board was finally 
released, no defects were found in the driver or the simulator, while the errors and 
not agreed changes introduced in the board where discovered during the first execution.

@@

In section \ref{sec:test_cases_on_the_job} the execution of the test suites
is described, the data collected is presented.

In section \ref{sec:test_cases_retrospectives} the data collected is analyzed, 
both from a qualitative and a quantitative points of view.

In section \ref{sec:test_cases_and_the_scenarios} the different scenarios on 
which test cases based on specifications can be applied are analyzed, and 
the resulting benefits it is possible to obtain are detailed.

@@

%=====================================================================
\section{Data stream and protocol description}
\label{sec:data_stream_and_protocol_description}

@@

%=====================================================================
\section{Data stream and protocol specification}
\label{sec:data_stream_and_protocol_specification}

@@

%=====================================================================
\section{Test cases}
\label{sec:test_cases}

@@

The dual testing approach was thought as a winning option, when 
compared to hand based tests or completely automated tests used alone, 
testing a driver for a stream communication protocol with hard constraints 
on development time.
\ST stream communication protocol is characterized by a well definite 
packet structure, and a well definite semantic on the packet contents 
as well.
On the other hand in the protocol there are not strong constraints on 
synchronization, since the communication channel is splitted in two (in and 
out) separate concurrent and asynchronous channels.
The lack of hard synchronization constraints reduce the advantage to use 
a classical state machine based specification \todo{references: 
protocol tests generated from a state machine based specification}.

@@

\todo{describe the necessity to add specification generated tests: to ensure
that specification is sound, to obtain higher coverage}

@@

To have proper unit test suite, for each method under test a proper stub, 
driver and oracle should be identified, these are the usual components found 
in the harness that enable test execution\cite{binder1999testing}.
In JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, the framework used 
to manage the implementation and execution of hand made and generated unit 
tests, there is no clear distinction between drivers, stubs, and oracle of a 
test; noneless, even if a proper distinction is not imposed by the framework, 
the conceptual distinction still holds.

@@

\todo{insert references to testing embedded systems with simulators and/or 
with driver/stubs}

\todo{describe approach: high level informal specifications with BON, 
iterative refinements of code, to distill test cases and formal 
specification of the interfaces} \notev{remove, BON has not been used 
properly, maybe a follow up paper}

%=====================================================================
\section{Test cases on the job}
\label{sec:test_cases_on_the_job}

@@

The first \STSB prototype delivered was meant to build empty packets,
but the packets were built in a way not conformant to the
specification, the sentinel values used to delimit a packet were not
correct.  The board implementation error has been detected by the
\lil{StreamDriver} package.

@@

The formal model and the run time assertion checker enabled the
detection of errors on the \SB itself.  Errors regarding the expected
values and values ranges of the filds contained in the packets and the
frames (4 protocol errors), and errors regarding the dependencies
between values in field (2 protocol errors) were detected. No other
errors, other than the ones immediately discovered by the run time
assertion checker were later on discovered.

The hand made test suite is composed by over 140 tests, while the
generated one is 100 times bigger, totalling over 14000 tests.

@@

To compare in a quantitative way the effectiveness of the test suites,
code coverage metrics have been used; this is not the only measure
that have been considered, since it is almost unanimously accepted
that code coverage alone cannot asses the quality of a test suite
\todo{reference to: ``How to misuse code coverage'', ??? ``Coverage
  Metrics for Formal Verification''}.

@@

%=====================================================================
\section{Retrospective on test cases effectiveness}
\label{sec:test_cases_retrospectives}

@@

The statement code coverage figures shows that the test suites do a
good job on the most important package, that is, the
\lil{StreamDriver}. The generated test suit reaching almost $ 80 \% $,
while the hand made tests have only a slightly lower coverage ratio: $
76 \% $. But the combination of test suites is interesting,
raising the coverage ratio to $ 88.5 \% $. This is a clear indication
that one test suite is not completely overlapping the other; that is,
the test suites are not to be considered alternatives, but they have
to be used together to achive the higher benefits, they are
complementary. This consideration can be confirmed looking at the
figures obtained for \lil{SimulatorDriver} package as well.

The qualitative difference of the two test suites can be understood
analysing how the suites are built. The hand made test suite primarly
focuses on small features and use cases fragments that the system must
implement, or for which an exceptional case has been considered. A
test case is to be considered a rough specification written in
imperative style. The common structure require an object
initialization, followed by a hopefully not too long sequence of
public methods to bring one object into a desired state, than the
method under test is called and the object state is finally inspected
to verify it is in the expected state. The generated tests have a far
more limited scope, since only one method call is performed. Thus it
is difficult to explore complex behaviour. It is true that with a
complete specification, and specifing large enough data ranges types,
the genrated tests could teoretically achive the same expressive power
as the hand made ones, but the price to pay would be expensive: much
longer an more complex specifications, requiring more time to be
designed and or updated, and more tests to be run, augmentig the
already not so short computation time of the generated test suite.

\todo{compare qualitative: hand made, many calls, vs generated, single call, tests}

\todo{compare compare coupled approach vs single approaches}

\todo{comment quantitative: number of bugs found}

\todo{compare qualitative: spec driven dev, test driven dev, test
  oriented code is an automatic result}

\todo{spec and hand useful to drive, spec and hand and simulator
  useful to test, spec and generated useful for statement coverage and
  easy methods, spec useful to test hardware, hand useful to spec,
  spec useful to hand... test driven spec}

%=====================================================================
\section{Test cases and the scenarios}
\label{sec:test_cases_and_the_scenarios}

\todo{benefit: the approach can be used to test different part of the system
that share the same specifications}

\todo{testing the protocols layers}

\todo{testing the simulators}

\todo{describe general approach: depending at what level (of abstraction) tests 
and specifications are written, it is possible to verify both the simulator and 
the real protocol (isn't this obvious?)}

%=====================================================================
\section{Related works}
\label{sec:related_works}

%=====================================================================
\section{Conclusion}
\label{sec:conclusion}

%======================================================================
%% \nocite{ex1,ex2}
\bibliographystyle{latex8}
\bibliography{extra,%
              abbrev,%
              ads,%
              category,%
              complexity,%
              hypertext,%
              icsr,%
              knowledge,%
              languages,%
              linguistics,%
              meta,%
              metrics,%
              misc,%
              modeling,%
              modeltheory,%
              reuse,%
              rewriting,%
              softeng,%
              specification,%
              ssr,%
              technology,%
              theory,%
              web,%
              upcoming,%
              upcoming_conferences,%
              conferences,%
              workshops,%
              verification,%
              escjava,%
              jml,%
              nijmegen}

%======================================================================
% Fin

>>>>

% \newcommand{\notev}[1]{\xspace$\textcolor{blue}{\omega^\textsf{vieri}}$\marginpar{\scriptsize\textsf{Vieri:}
%     #1}}
\newcommand{\notev}[1]{\xspace\marginpar{\scriptsize\textsf{Vieri:}
    #1}}
\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:}
    #1}}
\newcommand{\noted}[1]{\xspace$\textcolor{red}{\dagger^\textsf{dragan}}$\marginpar{\scriptsize\textsf{Dragan:}
    #1}}

@@

% ---------------------------------------------------------------------
% New commands, macros, \etc
% ---------------------------------------------------------------------

@@

% =====================================================================

@@

\author{Joseph R. Kiniry, Vieri del Bianco and Dragan Stosic\\
  UCD CASL: Complex and Adaptive Systems Laboratory and\\
  School of Computer Science and Informatics,\\
  University College Dublin,\\
  Belfield, Dublin 4, Ireland,\\
  kiniry@acm.org, vieri.delbianco@ucd.ie and dragan.stosic@gmail.com\\
}

@@

% ======================================================================
% \thispagestyle{empty}
\begin{abstract}

  Sensor equipped devices typically communicate with a streaming
  packet-based network protocol.  The communication is focused on
  controlling the sensors and report the sensors measurements.  The
  \STSB is an example of such a device, it communicates with an
  arbitrary computer system, the \STPU over a USB bus using a
  streaming packet-based protocol.  The \STU is a custom-designed, but
  inexpensive, sensor platform: it is composed by a one or more \STSBs
  paired with a general-purpose small-scale compute node, the \STPU
  (including everything from PDAs to powerful touchscreen-based
  portable computers).  There are over a dozen sensors on the \STSB
  itself, and new sensors can be added either using expansion slots on
  the \STSB, or adding them directly to the USB bus.  A third-party
  low-level library, provided by the USB chip vendor, the USB chip
  that has been used in the \STSB, creates and parses generic USB
  communication packets.

  A custom medium-level Java driver creates and parses the \STSB
  streaming communication packets.  The driver is formally specified,
  validated and verified with rigorous unit testing, both hand made
  and automatically generated.  Because of this process and
  combination of technologies, using this specification we are able to
  both verify a real \ST and a simulated one.  The system have been
  developed concurrently with the protocol specification and the real
  \STSB, the board were not available until the last months of
  development.  The protocol model is itself an executable stream
  simulator that can be used to develop new \ST user applications
  without having the \ST hardware at all.  Consequently, from this
  single specification, we can test that the \ST hardware communicates
  as expected, test that our \ST simulator communicates as expected,
  and simulate the \ST protocol stream itself.

@@

% ======================================================================
\section{Introduction}
\label{sec:introduction}

@@

Because of hard time constraints, and the initial unavailability of
the custom board, we needed to progress concurrently with all the
development tasks: the specification of the communication protocol
with the \STSB, the development of the controlling code in the \STSB,
the development of the driver to communicate with the \STSB, and the
development of the \STSB simulators needed for future tests of the
system (the board itself with the embedded software controlling it was
developed by an external manufacturer, hence, it is not directly taken
into account here).

There exist various approaches in literature addressing this kind of
development constraints (rapid development in hardware software
codesign), but they are all focused on big scale systems.  The
hardware software formal codesign methodologies usually have a common
structure\cite{Slomka2000,Hoffman2001}:

\begin{itemize}
\item High level specification of the system.
\item Translation of the high level specification to lower level
  specifications: hardware and software counterparts.
\item Generation of software code.
\item Run hardware specification into a simulator.
\item Build hardware component.
\end{itemize}

Even when cosidering the development of the device hardware, embedded
software and software driver as separate entities that possibly have
to be developed concurrently, the approaches proposed are similar in
structure to the one seen in the case of hardware software
codesign\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus
on very accurate formal specification that drive all the other
development tasks.

This approach is not feasible or convenient in our case: we want to
develop all the software components concurrently, we want to enhance
the protocol specification incrementally; we are not able to provide a
complete specification from the beginning, we are not interested in
direct code generation from specification, we want to generate
simulators to be used to feed the remaining system, even if the \STSB
is not available, and we want to be able to verify the \STSB behavior
when it is available.

To tackle the complexity of the concurrent tasks the formal model of
the protocol was used as a focus point that led all the other
development tasks (in this part, the approach is similar to the one
used in rapid development in hardware software codesign, as mentioned
above).  The protocol to specify has been divided in various (thin)
layers, each of the layer has been specified with a different
(lighweight) approach, some of the layers are completely formalized.
The formal model of the protocol was used to check at run time the
correctness of the /SB, the driver, the simulator, and the hand made
test suites of the driver and the simulator.  The formal model was
used to automatically generate a test suite to verify that both the
driver and the simulator were compliant to the specification, and
eventually to verify the specification itself.

The approach proved itself effective, at the point that, when the
board was finally released, only one small defect were found in the
driver or the simulator, while the errors and not agreed changes
introduced in the board where discovered with no effort.  Besides, the
simulators were able to feed with correct data the rest of the system

@@

In section \ref{sec:test_cases_on_the_job} the execution of the test
suites is described, the results are presented.

In section \ref{sec:test_cases_retrospectives} the data collected and
the results are analyzed, both from a qualitative and a quantitative
points of view.

@@

% =====================================================================
\section{Data stream and protocol description}
\label{sec:data_stream_and_protocol_description}

@@

% =====================================================================
\section{Data stream and protocol specification}
\label{sec:data_stream_and_protocol_specification}

@@

% =====================================================================
\section{Test cases}
\label{sec:test_cases}

@@

The dual testing approach was thought as a winning option, when
compared to hand based tests or completely automated tests used alone,
testing a driver for a stream communication protocol with hard
constraints on development time.  \ST stream communication protocol is
characterized by a well definite packet structure, and a well definite
semantic on the packet contents as well.  On the other hand in the
protocol there are not strong constraints on synchronization, since
the communication channel is splitted in two (in and out) separate
concurrent and asynchronous channels.  The lack of hard
synchronization constraints reduce the advantage to use a classical
state machine based specification (as the one presented in
\cite{Hubbers2004}).

@@

@@

To have proper unit test suite, for each method under test a proper
stub, driver and oracle should be identified, these are the usual
components found in the harness that enable test
execution\cite{Binder1999}.  In
JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, the framework
used to manage the implementation and execution of hand made and
generated unit tests, there is no clear distinction between drivers,
stubs, and oracle of a test; none-less, even if a proper distinction
is not imposed by the framework, the conceptual distinction still
holds.

@@

% =====================================================================
\section{Test cases on the job}
\label{sec:test_cases_on_the_job}

@@

The first \STSB prototype delivered was meant to build empty packets,
but the packets were built in a way not consistent with the
specification, the sentinel values used to delimit a packet were not
correct.  The board implementation error has been detected by the
\lil{StreamDriver} package.

@@

The formal model and the run time assertion checker enabled the
detection of errors on the \SB itself.  Errors regarding the expected
values and values ranges of the fields contained in the packets and
the frames (4 protocol errors), and errors regarding the dependencies
between values in field (2 protocol errors) were detected. No other
errors, other than the ones immediately discovered by the run time
assertion checker were later on discovered.

The hand made test suite is composed by over 140 tests, while the
generated one is 100 times bigger, totaling over 14000 tests.

@@

To compare in a quantitative way the effectiveness of the test suites,
code coverage metrics have been used; this is not the only measure
that have been considered, since it is almost unanimously accepted
that code coverage alone cannot asses the quality of a test
suite\cite{Marick1999,Chockler2006}.

@@

Execution time are very different, the hand made test suite, with the
run time assertion checking active runs, completely in less the one
minute on an average PC, while the generated test suite runs in more
than 20 minutes (the rough ratio is $ 1 : 65 $). Most of the time is
spent on setting up the suite, only considering the real execution of
the tests, with no suite setup, the execution time reduces to less
than 4 minutes (the rough ratio is $ 1 : 10 $).

% =====================================================================
\section{Retrospective on test cases effectiveness}
\label{sec:test_cases_retrospectives}

@@

The statement code coverage figures shows that the test suites do a
good job on the most important package, that is, the
\lil{StreamDriver}. The generated test suit reaching almost $ 80 \% $,
while the hand made tests have only a slightly lower coverage ratio: $
76 \% $. But the combination of test suites is interesting, raising
the coverage ratio to $ 88.5 \% $. This is a clear indication that one
test suite is not completely overlapping the other; that is, the test
suites are not to be considered alternatives, but they have to be used
together to achieve the higher benefits, they are complementary. This
consideration can be confirmed looking at the figures obtained for
\lil{SimulatorDriver} package as well.

The qualitative difference of the two test suites can be understood
analyzing how the suites are built. The hand made test suite primarily
focuses on small features and use cases fragments that the system must
implement, or for which an exceptional case has been specified. A test
case is to be considered a rough specification written in imperative
style. The common structure require an object initialization, followed
by a hopefully not too long sequence of public methods to bring one
object (or set of objects) into a desired state, than the method under
test is called and the object state is finally inspected to verify it
is in the expected state. Mock objects can be used to limit the
dependencies. The test suite should be composed only by unit tests,
but usually some tests are more complex to be called proper unit
tests, they can be seen as simple integration tests.

The generated test suite has a more focused and limited scope, since
only one method call is performed in each test, hence it is difficult
to explore complex behaviors. It is true that with a complete
specification, and specifying large enough data ranges (the
appropariate data ranges are combined to generate the tests), the
generated tests could theoretically achieve the same expressive power
as the hand made ones, but the price to pay would be expensive: much
longer and more complex specifications, requiring more time to be
designed and or updated, and more tests to be run, augmenting the
already not so short computation time of the generated test suite.

The test suites gave their best when used in a combined approach: the
generated test suite, checking most simple behaviors, while the hand
made checking the more complex ones.  During the development of the
driver code and the simulators, the generated test suit was helpful to
test the incrementally growing specification and the finer details of
each method, while the hand made one was helpful to verify the
behaviour on complex methods calls on objects intialized to specific
states. Having an hand made test suite also make the code style more
test oriented, which usually leads to a better quality and more
modularized source code\cite{Binder1999}.

The test suites has been used to verify different parts of the
systems, that is, the real system as well as the simulators.  The
multiple simulators and multiple test suites approach is not new, it
is often used in embedded systems software testing\cite{Broekman2002}.
This is a general approach that can (and should) be used to verify
both the simulators built and the real systems.  This approach can be
easily used only if the code has been structured following the already
mentioned dependency inversion principle\cite{martin1996dependency}
(see Figure \ref{fig:class_diagram_main}).

The combined test approach, driven by the packet protocol
specification, prooved succesfull on verifing the protocol itself and
the software implementing it.  Neverthless, it has to be noted that we
did not verified the highest levels of the protocol (see
\ref{sec:data_stream_and_protocol_description}: ``Packet sequence
rules'' and ``Packet input output asynchronous rules'').  Since on
these levels the protocol properties regards sequences of packets,
testing them can be tricky, since the sequences of packets are not
referred to explicitly in the implementation.  Two possible strategies
can be used:

\begin{itemize}
\item Translate the packet sequence properties in local packet
  properties (when this is possible).  The advantage is to have
  properties shown as the usual precondition, post conditions and
  invariants.  But the localized properties can be more complex and
  less intuitive and comprehensible than the original ones; finally,
  some of the global properties could proove themselves impossible to
  be translated into local ones.
\item Use separate utility classes, meant to verify a sequence of
  packages.  This approach is not clean, since it requires adding and
  using new classes, unneeded by the core implementation, to verify
  the protocol.  It has the advantage of keeping the properties that
  have to be verified simple.
\end{itemize}

One problem we noticed regards the executiuon time of the generated
tests.  The executiuon time is long for two reasons:

\begin{itemize}
\item Number of the tests generated. This problem is not directly
  addressable, since it is the heart of the test suite generation
  creation method itself: generating all the combinations of possible
  inputs for each method under test.
\item Runtime assertion checker. Each test is executed while RAC
  checks all the relavant properties. This problem, again, is not
  directly addressable. But, we can speculate that if the RAC
  assertions could be triggered on and off on single properties in a
  dynamic way, that is, during runtime, a test suite could use take
  advantage of this possibility, turning on the RAC only on the
  properties that have to be verified for the classes or methods under
  tests (and not for all the involved classes and methods); this could
  save time during the test suite execution.
\end{itemize}

% =====================================================================
\section{Related works}
\label{sec:related_works}

JML has been used previously to verify a communication protocol in
\cite{Hubbers2004}, but the protocol is a smart card protocol, and it
is well represented through state machines, which are then refined
into JML properties.  The approach is in waterfall style, the need to
develop the protocol, the specification and the implementation
concurrently is not addressed.  No hand made or generated test suites
are used, the code is specified with JML assertions, refined by the
state machine specifications, and then checked statically with
ESC/Java2 static checker\cite{CokKiniry04} and at runtime with JML2
tools RAC\cite{BurdyEtal05-STTT}.

The JML-JUnit automatic test suite generation tool has not been widely
used or assessed. It is described by its own creators in
\cite{Cheon2002,Cheon2004,Cheon2005}, it has been used in
\cite{Oriat2004} and \cite{Cheon2005}, but only as a stand alone test
suite, and, more importantly, only against proof of concepts system,
not real complex systems.  A brief assessment of the JML-JUnit tool
has been made in \cite{Tan2004}, the analysis is based on injecting
random generated errors in the source code, and then evaluating the
ratio of the errors found by the test suite: the results shows that
the JML-JUnit tool is not effective enough if used as a stand alone
tool. No attempts on running or evaluating the generated test suite
alongside with an hand made test suite are made.

% =====================================================================
\section{Conclusion}
\label{sec:conclusion}

\todo{specification and hand tests useful to drive, specification and
  hand tests and simulator useful to test, specification and generated
  tests useful for statement coverage and easy methods, specification
  useful to test hardware, hand useful to spec, spec useful to
  hand... test driven spec}

\section{Next papers}

\todo{remove and push this on a new paper}

\paragraph*{TDS: Test Driven Specification}

Approach:
\begin{itemize}
\item more specification used together: formal and not formal
\item formal specification is not complete: it is just good
  ``enough'', static checkers unfortunately cannot be used, since the
  specification is partial
\item use runtime asserion checker for formal specifications, on tests
\item use generated tests for formal specifications
\end{itemize}

Who drives what:
\begin{itemize}
\item specification (jml - formal specification or unit test -
  examples fragments) driven development
\item test driven specification
\item formal specification is not complete: it is just good
  ``enough'', static checkers unfortunately cannot be used, since the
  specification is partial
\item use runtime asserion checker for formal specifications, on tests
\item use generated tests for formal specifications
\end{itemize}

Benefits:
\begin{itemize}
\item real agile approach
\item real formal approach
\item use best of both worlds, do not try to use only on of them
\end{itemize}

Case example: actual paper, modified to fit the new parts, expecially
the test driven specification part.

\todo{describe approach: high level informal specifications with BON,
  iterative refinements of code, to distill test cases and formal
  specification of the interfaces}

% ======================================================================
%% \nocite{ex1,ex2}
\bibliographystyle{latex8} \bibliography{extra,%
  abbrev,%
  ads,%
  category,%
  complexity,%
  hypertext,%
  icsr,%
  knowledge,%
  languages,%
  linguistics,%
  meta,%
  metrics,%
  misc,%
  modeling,%
  modeltheory,%
  reuse,%
  rewriting,%
  softeng,%
  specification,%
  ssr,%
  technology,%
  theory,%
  web,%
  upcoming,%
  upcoming_conferences,%
  conferences,%
  workshops,%
  verification,%
  escjava,%
  jml,%
  nijmegen}

% ======================================================================
% Fin
